---
title: "FlightDelay_Assignment"
author: "Haroon"
date: "July 15, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read File

```{r cars}


FlightData <- read.csv("C:\\MOOC\\Machine Learning\\AssociationRules\\FlightDelays.csv")

```

#Look at the summary of all the variables and convert the following variables as factors -> Weather + DAY_WEEK + Flight Status# 

```{r pressure, echo=FALSE}

summary(FlightData)
head(FlightData)
str(FlightData)
FlightData$Weather <- as.factor(FlightData$Weather)
FlightData$DAY_WEEK <- as.factor(FlightData$DAY_WEEK)
FlightData$Flight.Status <- as.factor(FlightData$Flight.Status)


```

#Convert the data frame in a transactions object. Look at the first 6 transactions to understand how 'arules' library requires data to be transformed. Use the below R code to view the transactions data

```{r}

FlightData$levels<-ifelse(FlightData$CRS_DEP_TIME>=600 & FlightData$CRS_DEP_TIME<=1200,"level1",
                      ifelse(FlightData$CRS_DEP_TIME>=1200 & FlightData$CRS_DEP_TIME<=1800,"level2",
                        ifelse(FlightData$CRS_DEP_TIME>=1800 & FlightData$CRS_DEP_TIME<=2100,"level3","level4")))

FlightData$levels <- as.factor(FlightData$levels)


```

```{r}

str(FlightData)

```

#Apply 'arules' algorithm and play with various support, lift and confidence values.

```{r}
  FlightData$CRS_DEP_TIME <- NULL  

  library(arules)
  rules <- apriori(FlightData,parameter = list(sup = 0.2, conf = 0.6,target="rules"))

```


#Inspect all the rules.

```{r}

summary(rules)
inspect(rules)

```


#Filter the rules with specific LHS and RHS conditions E.g.; Filter the rules with Flighstatus=0

```{r}

rules.itemfilter2 <- as(subset(rules, subset = lhs %in% 
                        "Flight.Status=0"),
                        "data.frame")
rules.itemfilter2 

rules_Lift <- as(subset(rules, subset = rhs %in% "DEST=LGA" & lift > 1.01),
                        "data.frame")
rules_Lift


```


#Filter redundant rules if any

```{r}

inspect(rules[is.redundant(rules)])
rules <- (rules[!is.redundant(rules)])

```



#Pot and visualize the rules

```{r}
install.packages('arulesViz')
library(arulesViz)
plot(rules)

```


#Write the Rules into CSV


```{r}

write(rules, file = "data.csv", sep = ",")

```

